\thetitle

### 2013-04-18 tfpie/tfpie.tex
finish conclusion

We have now completed the third iteration of this course, teaching
approximately 35 students each time.  Our experience has been
uniformly positive, and the students have gone on to significant
success in the subsequent courses, despite the curriculum differing
from what the bulk of Northeastern University computer science majors
take.

Perspective into the upper-class curriculum:

\begin{itemize}
\item types, contract, invariants : formal methods, program verification,

\item representation independence : component-based software engineering,
  data-structures, algorithms

\item interfaces, interactions, distribution : operating systems, networking
\end{itemize}

Other steps:
\begin{itemize}
\item types in class language
\item pedagogical Java environment
\item ugly syntax for class language
\item world in Java
\end{itemize}

\cite{dvanhorn:sicp}

\cite{local:htdc}

>>>>

We have now completed the third iteration of this course, teaching
approximately 35 students each time.  Our experience has been
uniformly positive, and the students have gone on to significant
success in the subsequent courses, despite the curriculum differing
from what the bulk of Northeastern University computer science majors
take.  The class has also had notable success in the recruitment and
retention of female students, as compared to the other versions of the
second-semester course.

The course has also provided a vantage point to introduce topics that
will be taken up later in the curriculum.  We present types,
contracts, invariants, and properties of functions, all of which tie
into both the concurrent course on logic and computation, as well as
later classes on formal methods.  The emphasis on
representation-independence and interfaces  both ties into later
classes on software engineering, as well as preparing the ground for
algorithms and data structures courses.  Finally, the use of
interactive and distributed systems connects to later courses on
operating systems and networks.

Despite our success, much remains to be done.  Type systems are a
fundamental concept, but their introduction accompanies the rest of
Java.  Developing a typed version of our introductory languages would
allow a smoother introduction of this idea.

Our class's use of Eclipse could also be improved by first
transitioning to a pedagogically-oriented Java environment, but we
have not evaluated the specific options.  Alternatively, introducing
Java-like syntax for the teaching languages we have developed would
help tease apart the difficult transitions still present in the course.

Finally, the Java portion of the class does not continue the use of
``World''-style interactive graphical
programming. \citet{local:java-world}.  Instead, our course focuses on
coverage of standard Java libraries, as well as introductory
algorithmic and data structure topics.  Continuing to use World-style
programming in motivating examples might be valuable for continutity
between the two halves of the course.
